### 2023-11-27 draft_jss/draft_jss.Rmd
update draft jss

## Reservoir networks presentation

Reservoir Computing (RC) is a prominent machine learning methodology
that has gained significant attention in recent years for its ability to
effectively process information generated by dynamical systems. This
innovative approach leverages the dynamics of a high-dimensional
reservoir to perform complex computations and solve various tasks based
on the response of this dynamical system to input signals. Reservoir
Computing is characterized by its unique architecture, which includes a
dynamic system, often referred to as a reservoir, that projects temporal
input signals onto a high-dimensional feature space. This
high-dimensional representation of input data enables Reservoir
Computing to excel in tasks such as time series prediction, pattern
recognition, and signal processing.

**TODO : cite : Jaeger, Herbert. “The “echo state” approach to analysing and training recurrent neural networks-with an erratum note.” Bonn, Germany: German National Research Center for Information Technology GMD Technical Report 148.34 (2001): 13.**
**Maass W, Natschläger T, Markram H. Real-time computing without stable states: a new framework for neural computation based on perturbations. Neural Comput. 2002;14(11):2531‐2560.**

## Why reservoir are interesting compared to standard approaches (penalised regression) and deep learning approaches?

RC exhibits several advantages over its counterparts. It excels in tasks
that require the processing of temporal data, such as time series
forecasting and speech recognition, thanks to its inherent memory
capabilities. RC offers computational efficiency and reduced susceptibility
to overfitting compared to deep learning. Moreover, RC is able to capture intricate temporal
patterns compared to statistical methods such as penalized regression.

**TODO : find ref to support it.**

## examples of previous use

**TODO: add ref from use case intro on use in epidemiology + find review**
**Table 1 contains a list of use case of RC : https://www.sciencedirect.com/science/article/pii/S0893608019300784**

## challenges of using reservoir

RC use is still challenging for several reasons. First, its stochastic
nature due to the reservoir random part makes the results of a given RC
unpredictable even if the hyperparameters are well chosen. Second,
even though, RC relies on several hyperparameter to determine the
connections between neurons. There exist few guidance about the way to
set the hyperparameters and most often the user must rely on a wrapper
approach were combinations of hyperparameters performance are evaluated
on a train set before choosing the ideal combination one. Third, most
published paper are based on a low dimensional setting and few guidance
exist in the context of high dimensional data, especially considering
the choice of hyperparameters. Fourth, the use of reservoir computing in
an epidemic setting with highly non-stationary time series is scarce
whereas it is of tremendous interest for epidemiologists. Fifth, there
is currently no implementation available on R making the use of this
method challenging for user not familiar with Python.

## what we do

In this paper we propose to explore those challenges with extensive
guidance in order to help new user to fully benefit from RC. First a
general introduction to reservoir computing is provided with a tutorial
for its use using `reservoirnet`, a R package based on `reservoirPy` Python
module. Second, we explore the different challenges of the use of RC in
a non-stationary high dimensional setting based on the use case of
covid-19 hospitalizations forecast with extensive guidance on the
modelling strategy, the choice of hyperparameters using a genetic
algorithm and the implementation.

@@

The parameter $\lambda$ is the ridge penalization which aims to prevent
overfitting.

Overall, there are four main hyperparameters to be chosen by the user: the leaking rate which defines the memory of the RC, the input scaling which define the relative importance of the features, the spectral radius which define the connections of the neurons inside the reservoir which in turn define the degree of non-linear combination of features and the ridge penalization which controls the degree of overfitting. The choice of hyperparameter is difficult and the use of a wrapper approach where the performance of the RC with different combinations of hyperparameters is evaluated on a train set, the best combination is chosen for the use on the test set.

@@

We can proceed by loading useful packages (i.e ggplot2 [@wickham_ggplot2_2016] and dplyr [@wickham_dplyr_2023], **TODO : add cowplot**), data and define the task:

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(reservoirnet)
library(cowplot)
library(patchwork)

@@

`TODO :` check for new published forecast models

RC can be approached as an extension of penalized linear regression where inputs are processed by a reservoir which allows memory and non-linear combinations. As penalized linear regression was the best approach for covid-19 forecast in @ferte_benefit_2022 and that RC showed promising results for epidemic forecast in @ghosh_reservoir_2021, @liu_nanophotonic_2023 and @ray_optimized_2021, we proposed to use RC to forecast hospitalizations at 14 days at the University Hospital of Bordeaux. The main objective of this study is to evaluate the performance of RC for this task. Secondary objective will be (i) to assess the performance of RC depending on the input dimension, (ii) to assess the relevance of Genetic Algorithm (GA) for hyperparameter optimization and feature selection in this context.

@@

First and second derivative over the last $7$ days were computed to enrich model information. To take into account measurement error and daily noise variation, data were smoothed using a local polynomial regression with a span of 21 days.

@@

The task was to forecast 14 days ahead the number of hospitalized patients. As seen at section [Regression](#basicregression), we will train the model to predict the variation of hospitalization defined as $outcome_{t+14} = hosp_{t+14} - hosp_{t}$. Metrics computation and visualizations will be performed on the predicted number of hospitalizations denoted as $\hat{y_{t+14}} = \widehat{outcome_{t+14}} + hospitalizations_{t}$ and the observed number of hospitalizations at time $t+14$ denoted as $y_{t+14} = hospitalizations_{t+14}$

The dataset was separated into two periods. First period from May 16, 2020 to March 1, 2021 served to identify relevant hyperparameters. Remaining data was used to evaluate the model performance. As the time series is not stationary and most appropriate hyperparameter value might change over time, we also investigated the effect of hyperparameter update every month.

The performance of the model was evaluated according to several metrics: the median absolute error ($MAE = median(|\hat{y_t}-y_t|)$) , the median relative error ($MRE = median(|\frac{\hat{y_t}-y_t}{y_t}|)$), the median absolute error to baseline ($MAEB =  median(|\hat{y_t}-y_t| - |y_{t-14} - y_t|)$), the median relative error to baseline ($MREB = median(\frac{|\hat{y_t}-y_t|}{|y_{t-14} - y_t|})$). Such that the model denoted as baseline was the model defined as $\hat{y_t} = y_{t-14}$. All those metrics reflects different information. $MAE$ and $MAEB$ are highly sensitive to forecast error when the number of hospitalizations is high. $MAEB$ encapsulates information about the current dynamic of the hospitalizations : for the same absolute error, absolute error to baseline will be lower if the number of hospitalization is not constant. **MAEB = redundant with MAE ?**. $MRE$ is highly sensitive to forecast error when the number of hospitalizations is low because the denominator will be small. $MREB$ is highly sensitive to small forecast error when the number of hospitalized patients is constant over time. For each metric, the median was chosen over the mean to be robust to outliers. For instance, for relative error, when the outcome is low (e.g equal to 1 hospitalization), an error of 5 hospitalizations could results to a 400% relative error. **But that might favor models with chaotic metric because extreme values won't be used to compute the error ? => keep median for relative error and mean of absolute error and optimize on MAE instead of MREB ? Also, because MAE is a difference it is similar if we compute it on the number of hospitalization or on the variation of the number of hospitalization. Also we might discuss to penalize more the error when the number of hospitalizations is high (already done implicitly by AE) and when there is a dynamic shift => ponderate AE by absolute value of second derivative ?**.

### Feature selection with RC

We investigated the performance of RC using various feature selection approach and input scaling configuration. The following setting were evaluated :

- Feature selection based on expert inputs with common input scaling
- Feature selection based on expert inputs with specific input scaling for each feature
- Genetic algorithm feature selection with common input scaling
- Genetic algorithm feature selection with specific input scaling for each feature
- Elastic-net feature selection with common input scaling

Experts inputs were a priori defined features that were judge important for epidemic forecast : hospitalizations, positive RT-PCR, positive RT-PCR among the 60 years old and more, fraction of positive RT-PCR, fraction of positive RT-PCR among the 60 years old and more, IPTCC, number of patient EHR mentioning "covid-19" in emergency units department, number of people having received at least one dose of vaccine in Gironde. Additionally we provided the first derivative at 7 days for each feature.

We used feature selection using elastic-net because it was the best model in previous experiment compared to other machine learning (@ferte_benefit_2022). This feature selection impose to select the penalty hyperparameter of elastic-net. This hyperparameter was selected using genetic algorithm as depicted [previously](#geneticalgo).

Genetic algorithm feature selection was performed using a binary hyperparameter for each feature determining if it was to be selected or not.

In addition, we investigated if monthly update of hyperparameter was improving the performance and if optimizing the random seed (which defines the connections inside the reservoir) was improving the results.

### Hyperparameter optimisation using random search {#geneticalgo}

RC relies mainly on 4 hyperparameters including the leaking rate (i.e "memory" parameter), spectral radius (i.e "chaoticity" parameter), input scaling (i.e "feature gain" parameter) and ridge (i.e penalization parameter). Input scaling can be either, common to all features or specific to each feature which increases the number of hyperparameter by the number of features. As the relevancy of feature for forecast evolves over time, we explored (i) filter based feature selection based on elastic-net before RC which has a penalization hyperparameter denoted as $enet$ and (ii) wrapper feature selection for which each feature has an additional binary hyperparameter defining if it should be selected or not.

Regarding the number of hyperparameter, an automatic approach is necessary. We chose to use a genetic algorithm (GA) which has the benefit of making few assumptions on the link function between the vector of hyperparameter value and the objective function but might need many trials which are possible due to the fast of RC (**TODO : add ref**). The objective function of GA was to minimize the $MREB$ metric **(change ?)**. The pseudo-code of GA is defined as follow :

For each hyperparameter, we must define the boundaries of the search space. We chose a large search space because we had few hypothesis to restrict it. Spectral radius and input scaling were defined on $[10^{-5} ; 10^5]$, leaking rate on $[10^{-5} ; 1]$, ridge on $[10^{-10} ; 10^5]$.

The penalization for elastic-net feature selection was initially defined on $[10^{-10} ; 10^5]$. However, first experiments showed that, after few generations, GA was stuck in a area were no feature selection was performed and without the ability to sample new individuals with feature selection. This phenomena is likely to append for evolutionary algorithm where the sampling variance is small compared to the search space. This phenomena is depicted at figure \@ref(fig:enetNbFeatures) were there is no feature selection performed when penalization is below $10^{-4}$. To mitigate this problem, we chose to restrict the search space on $[10^{-4} ; 10^5]$.

@@

```{r}
ls_use_case <- readRDS(file = here::here("data/precomputed_results.rds"))
```

**TODO : change optimisation metric = mean absolute error ?**
**TODO : optimise the seed ?**
**TODO : compare with elastic-net alone ?**
**TODO : compare with LSTM ?**
**TODO : improve optimisation figures (too small)**
**TODO : do everything with the R API ?**
**TODO : provide clean replication code as a supplementary material (too long to be run inside a rmarkdown chunk)**

The goal of this task is to predict 14 days ahead the hospitalization curve depicted at figure \@ref(fig:casestudyhospitalizations). It shows both the training set (i.e before 2021-03-01) and the test set encompass several slope shifts which are difficult to forecast. We will first introduce the results of the genetic algorithm hyperparameter optimization and then present the forecast performance.

```{r ucPresentData, echo = FALSE, fig.cap = "Data of the use case. Outcome of interest is presented in orange. Model will be trained to forecast outcome variation but will be used to Outcome for performance and representation. Other features are represented in darkblue.", fig.height=10}
ls_use_case$plot_present_data
```

### Hyperparameter optimisation

**TODO : add final choice of hyperparameter value**

In this section we will first introduce the results for the features selected based on expert knowledge with multiple input scaling and for the features selected by genetic algorithm with multiple input scaling and seed optimization.

```{r ucHyperparam, fig.cap="Hyperparameter evaluation on training set by random search. Hp sets with MAE above 30 were removed for clarity of visualisation.", fig.height=12}
ls_use_case$plot_reservoir_hp / ls_use_case$plot_feature_input_scaling + plot_annotation(tag_levels = 'A')
```

@@

**TODO : if we keep this section, we should update with same hyperparameter (feature selection with only 30 features for GA multiple input scaling) and same aggregation mechanism as in the present forecast performance to have something clean.**
**Note : if we optimize the seed, then the repetition of a given set of hyperparameter makes no sense (i.e will provide same results each time).**

```{r ucAggregation, echo = FALSE, fig.cap = "Aggregation of several reservoir. Due to its intrisic chaoticity, a same hyperparameter set can results in heterogeneous forecast. Panel A shows the individual forecast of 40 repetition of the same hyperparameter set for the different RC configuration. Panel B shows the impact of aggregating several RC with the same set of HP. It has small interest when the input layer was not connected to output layer but significantly improves the performance otherwise.", fig.height=8}
ls_use_case$plot_before_aggregation_forecast + ls_use_case$plot_aggregation + plot_annotation(tag_levels = 'A')

### Present forecast performance

```{r perfMonthlyUpdate, echo = FALSE}
ls_use_case$table_perf_esn %>%
  knitr::kable(caption = "Model performance with several reservoir configuration. For each setting, 40 reservoir are computed and the forecast is the median of the 40 forecast. Results show the mean number of features and the performance metrics. MAE = Mean Absolute Error, MRE = Median Relative Error, MAEB = Mean Absolute Error to Baseline, MREB = Median Relative Error to Baseline.",
               col.names = c("Model", "MAE", "MRE", "MAEB", "MREB"),
               booktabs = TRUE,
               linesep = c('', '\\addlinespace'), digits = 2)
```

Table \@ref(tab:perfMonthlyUpdate) shows the performance on the test set. The number of selected features by elastic-net and by genetic algorithm were in the same order of magnitude. Updating the hyperparameter monthly did not always improve forecast accuracy (e.g epidemiological subset with multiple input scaling). Best model depend on the metric but epidemiological subset with multiple input scaling and no update was the best model or second to best model for all metrics.

Figure \@ref(fig:forecastwithorwithoutUpdate) shows the forecast with and without monthly hyperparameter update. Graphically we observe that the forecast of individual models (in grey) are chaotic and that forecast aggregation (in red) is necessary. This is less concerning for epidemiological subset with multiple input scaling or when the seed is controlled. In terms of performance, we observe similar conclusion as the one depicted for table \@ref(tab:perfMonthlyUpdate). However, the model errors did not occur for the same period for each model. For instance, the epidemiological with multiple scaling and monthly update error is located mainly during the winter 2022 increase. Model based on all features tend to overestimate the length of the summer 2021 wave.

@@

@@

>>>>

Reservoir Computing (RC) is a prominent machine learning methodology proposed by 
that has gained significant attention in recent years for its ability to
effectively process information generated by dynamical systems [@kim_health_2020, @maass_real-time_2002]. This
innovative approach leverages the dynamics of a high-dimensional
reservoir to perform complex computations and solve various tasks based
on the response of this dynamical system to input signals. RC has demonstrated its efficacy in tackling various challenges, encompassing pattern classification and time series forecasting in applications ranging from electrocardiogram analysis to bird calls, power plants, internet traffic, stock prices, and beyond [@tanaka_recent_2019].

RC is a kind of recurrent neural networks (RNNs) where the internal states, denoted as the reservoir, is randomly generated and solely the final layer is trained. The reservoir projects temporal input signals onto a high-dimensional feature space facilitating the learning of non-linear and temporal interactions. This offers the advantage of decreasing the computing time while consistently maintaining robust performance compared to conventional RNNs [@shahi_prediction_2022, @vlachas_backpropagation_2020].

The utilization of RC poses challenges on multiple fronts. Firstly, its stochastic nature, attributed to the random aspect of the reservoir, introduces unpredictability into the results, even with well-chosen hyperparameters. Secondly, RC relies on numerous hyperparameters to establish connections between neurons. Despite this, guidance on determining these hyperparameters is limited, often requiring users to adopt a wrapper approach where combinations of hyperparameter performance are assessed on a training set before selecting the optimal combination. Thirdly, there is a current absence of implementation in R, rendering the method challenging for users unfamiliar with Python [@trouvain_reservoirpy_2020, @steiner_pyrcn_2022] or Julia [@martinuzzi_reservoircomputingjl_2022].

In this paper, we endeavor to delve into these challenges, offering comprehensive guidance to assist new users in maximizing the benefits of RC. Initially, a broad introduction to reservoir computing is presented, accompanied by a tutorial on its application using `reservoirnet`, an R package built upon the `reservoirPy` Python module [@trouvain_reservoirpy_2020]. Subsequently, we investigate the various challenges associated with the use of RC for forecasting COVID-19 hospitalizations. This exploration includes detailed guidance on the modeling strategy, the selection of hyperparameters, and the implementation process.

@@

The parameter $\lambda$ is the ridge penalization which aims to prevent
overfitting. One can also connect the input layer to the output layer in addition to the reservoir nodes. In that case, $X$ is the accumulation of both such that :

$$X = \begin{bmatrix} x(1), u(1) \\ x(2), u(2) \\ ... \\ x(T), u(T) \end{bmatrix}
\text{ and } Y = \begin{bmatrix} y(1) \\ y(2) \\ ... \\ y(T) \end{bmatrix}$$

Overall, there are four main hyperparameters to be chosen by the user: the leaking rate which defines the memory of the RC, the input scaling which define the relative importance of the features, the spectral radius which define the connections of the neurons inside the reservoir which in turn define the degree of non-linear combination of features and the ridge penalization which controls the degree of overfitting. The choice of hyperparameter often rely on a wrapper approach where the performance of the RC with different combinations of hyperparameters is evaluated on a train set and the best combination is chosen to forecast on the test set.

@@

We can proceed by loading useful packages (i.e ggplot2 [@wickham_ggplot2_2016], dplyr [@wickham_dplyr_2023] and patchwork [@pedersen_patchwork_2023]), data and define the task:

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(patchwork)
library(reservoirnet)

@@

RC can be viewed as an extension of penalized linear regression, where inputs undergo processing by a reservoir, introducing the capacity for memory and non-linear combinations. Given the effectiveness of penalized linear regression in COVID-19 forecasting, as highlighted in @ferte_benefit_2022, and the promising results exhibited by RC in epidemic forecasting, as demonstrated in studies such as @ghosh_reservoir_2021, @liu_nanophotonic_2023, and @ray_optimized_2021, we have opted to employ RC for the prediction of hospitalizations at 14 days at the University Hospital of Bordeaux.

The primary aim of this study is to assess the performance of RC in this forecasting task. Secondary objectives include (i) comparing the performance of RC with that of elastic-net penalized regression (identified as the optimal model in @ferte_benefit_2022) and (ii) evaluating variations in RC performance based on different architectural choices, such as the connection between the input layer and the output layer, and the use of one input scaling per feature versus a common input scaling.

@@

First derivative over the last $7$ days were computed to enrich model information. To take into account measurement error and daily noise variation, data were smoothed using a local polynomial regression with a span of 21 days.

All data are publicly available. Weather data can be obtained from @smith_integrated_2011 using R package `worldmet` [@carslaw_worldmet_2023]. [Vaccine data](https://www.data.gouv.fr/fr/datasets/donnees-relatives-aux-personnes-vaccinees-contre-la-covid-19-1/) can be downloaded from @etalab_les_2020. EHRs data can be downloaded on dryad [@ferte_benefit_2023]. For privacy issues, publicly available EHRs data below 10 patients were obfuscated to 0. For convenience, all data were downloaded, merged and provided as replication material.

@@

The task was to forecast 14 days ahead the number of hospitalized patients. As seen at section [Regression](#basicregression), we will train the model to predict the variation of hospitalization, denoted as $hosp$, defined as $outcome_{t+14} = hosp_{t+14} - hosp_{t}$. Metrics computation and visualizations will be performed on the predicted number of hospitalizations denoted as $\widehat{hosp_{t+14}} = \widehat{outcome_{t+14}} + hosp_{t}$.

The dataset was separated into two periods. First period from May 16, 2020 to March 1, 2021 served to identify relevant hyperparameters. Remaining data was used to evaluate the model performance.

The performance of the model was evaluated according to several metrics:

- the mean absolute error : $MAE = mean(|\widehat{hosp_{t+14}}-hosp_{t+14}|)$.
- the median relative error : $MRE = median(|\frac{\widehat{hosp_{t+14}}-hosp_{t+14}}{hosp_{t+14}}|)$.
- the mean absolute error to baseline : $MAEB =  mean(|\widehat{hosp_{t+14}}-hosp_{t+14}| - |hosp_{t}-hosp_{t+14}|)$.
- the median relative error to baseline : $MREB = median(|\frac{\widehat{hosp_{t+14}}-hosp_{t+14}}{hosp_{t}-hosp_{t+14}}|)$)

Mean was chosen over median for $MRE$ and $MREB$ because those metrics tend to have extremely high values when the denominator is close to 0 (i.e when the number of hospitalized patients is close to 0 or the number of patients hospitalized at 14 days is close to the current number of hospitalized patients respectively). $MAEB$ and $MREB$ compare model performance to a baseline model which predicts the current number of hospitalized patients at 14 days. Those metrics help to determine the information added by the model and is a good baseline as covid-19 forecast model do not always outperform this basic forecast (@cramer_evaluation_2022).

### Models

We compared RC to elastic-net penalized regression (denoted as Enet). Furthermore we evaluated RC based on several architectures. First we compared RC with a single input scaling common to all features and a RC with on specific input scaling per feature. Second we compared RC where the input layer is connected to the output layer in addition to the connection between reservoir and output layer. Therefore, five models were evaluated :

- Elastic-net penalized regression denoted *Enet*
- RC with a single input scaling and no connection between input and ouput layers denoted *Common IS R %>>% O*
- RC with a single input scaling and connection between input and ouput layers denoted *Common IS I+R %>>% O*
- RC with multiple input scaling and no connection between input and ouput layers denoted *Multiple IS R %>>% O*
- RC with multiple input scaling and connection between input and ouput layers denoted *Multiple IS I+R %>>% O*

Because of the randomness of the reservoir, we took the median forecast of 20 reservoir on the train set and 40 reservoir on the test set. In addition, because covid-19 hospitalization is a non-stationary process, models were re-trained everyday using all previous days. To ease computation burden, only one day over two was used to find hyperparameters on the training set.

### Hyperparameter optimisation using random search

RC relies mainly on 4 hyperparameters including the leaking rate (i.e "memory" parameter), spectral radius (i.e "chaoticity" parameter), input scaling (i.e "feature gain" parameter) and ridge (i.e penalization parameter). Input scaling can be either, common to all features or specific to each feature which increases the number of hyperparameter by the number of features.

Following the notation from `glmnet` package [@friedman_regularization_2010], elastic-net penalized linear regression relies on two hyperparameters, lambda (i.e the penalization parameter) and alpha (i.e the compromise between lasso and ridge penalty)

Hyperparameter were selected in the training set (i.e before March 1, 2021) using a wrapper approach and a random search sampler using 2000 samples for each model. The sampling distribution were defined as follow :

- (RC) ridge and (Enet) lambda : log-uniform law defined between 1e-10 and 1e5
- (RC) input scaling and spectral radius : log-uniform law defined between 1e-5 and 1e5
- (RC) leaking rate : log-uniform law defined between 1e-3 and 1
- (Enet) alpha : uniform defined between 0 and 1

We provided large search space for all hyper-parameters. Search space was slightly reduced for leaking rate based on previous results and because a leaking rate of 1e-3 already imply that new inputs make the reservoir change really slowly which is not inline with the dynamic of covid-19 but would be appropriate for an application where the phenomena to forecast has a slow dynamic.

Finally, we provided an additional Enet model similar to the one in @ferte_benefit_2022 where alpha was set to $0.5$ and lambda was re-evaluated everyday in the test set based on previous data using the cross-validation procedure provided by `glmnet`.

@@

```{r echo=FALSE}
ls_use_case <- readRDS(file = here::here("data/precomputed_results.rds"))
```

The goal of this task is to predict 14 days ahead the hospitalization. Figure \@ref(fig:ucPresentData) shows both the training set (i.e before 2021-03-01) and the test set where the blue curve correspond to the input features (first derivatives are not shown) and the orange curves correspond to the outcome the model is trained on (i.e the hospitalization variation) and the hospitalizations at 14 days on which the performance metrics are computed. The figures outline that the relation between the input features and the outcome evolve over time and that the time series is not stationary. For instance IPTCC seems correlated to the outcome except that it completely miss the summer 2021 increase.

```{r ucPresentData, echo = FALSE, fig.cap = "Covid-19 epidemic at BUH. Outcome of interest is presented in orange. Model is trained to forecast Outcome curve which corresonds to the difference between Hospitalisatiosn at 14 days and current hospitalisations. Other features are represented in darkblue.", fig.height=8}
ls_use_case$plot_present_data
```

### Hyperparameter selection

Figure \@ref(fig:ucHyperparam) shows the hyperparameter optimisation using random search for the different RC architectures. We note that the multiple input scaling setting with input layer connected to output layer showed poor performance on the training set which might be due to an unextensive search of hyperparameter regarding the better performance of the common input scaling with connected input layer which is a specific configuration where all input scaling are equal.

As expected, we observe that the optimal leaking rate is close to 1 for all RC which is coherent with the short term dynamic of covid-19 epidemic. Trend for other hyperparameter are less clear even though best hyperparameter sets were close for RC with common input scaling.

```{r ucHyperparam, echo=FALSE, fig.cap="Hyperparameter evaluation on training set by random search. Hp sets with MAE above 30 were removed for clarity of visualisation.", fig.height=8}
ls_use_case$plot_reservoir_hp
```

Figure \@ref(fig:ucHyperparamMultipleIS) shows the hyperparameter search for RC with multiple input scaling and connected input layer. We observe that the random search tend to favor low importance given to vaccine and high importance given to derivative of hospitalisations and derivative of positive RT-PCR in the elderly. The rest of the feature do not show clear pattern. The lack of importance of number of people with one dose of vaccine in this setting was discussed by @ferte_benefit_2022 and related to the lack of vaccinated people in the training set as depicted at figure \@ref(fig:ucPresentData).

```{r ucHyperparamMultipleIS, echo=FALSE, fig.cap="Hyperparameter evaluation on training set by random search of the model with multiple input scaling and no connection between input layer and output layer. Hp sets with MAE above 30 were removed for clarity of visualisation.", fig.height=8}
ls_use_case$plot_feature_input_scaling
```

### Forecast performance

```{r performance, echo = FALSE}
ls_use_case$table_perf_esn %>%
  knitr::kable(caption = "Model performance with several reservoir configuration. For each setting, 40 reservoir are computed and the forecast is the median of the 40 forecast. Results show the performance metrics : MAE = Mean Absolute Error, MRE = Median Relative Error, MAEB = Mean Absolute Error to Baseline, MREB = Median Relative Error to Baseline.",
               col.names = c("Model", "MAE", "MRE", "MAEB", "MREB"),
               booktabs = TRUE,
               linesep = c('', '\\addlinespace'), digits = 2)
```

Table \@ref(tab:performance) shows the performance on the test set. Best model according to all metrics was RC with common input scaling and no connection between input and output layers. Having one input scaling per feature did not improve the model which might be due to low generalisability of the hyperparameter of the training set to the test set due to non-stationarity. Additionaly, connecting input layer to output layer did not improve the model, possibly because it add more dimension to the problem while the information provided is already inside the reservoir. We observe that elastic-net did not perform well and stay close to the baseline prediction. However, when the penalisation is updated each day, the peformance improve. This is emphasize the problem of the transferrability of the hyperparameters to which multiple input scaling models seem more sensitive as they have more hyperparameters. For instance, vaccine will keep a low weight whereas its low initial importance quickly increase to mitigate the effect of the high RT-PCR increase on the hospitalisation dynamic.

Figure \@ref(fig:forecastwithorwithoutUpdate) shows the forecast of the different models. We note that models struggle to accurately forecast slope shifts. For instance, summer 2021 initial increase is missed by all models, summer decrease is partially anticipated by RC with common IS and no input layer connection, winter 2021 increase is anticipated by all models but they tend to overestimate it because of the rise of vaccine effect.

@@

Figures \@ref(fig:ucAggregationSpread) and figure \@ref(fig:ucAggregationNb) show 40 model forecast for each of the best hyperparameter set of the different reservoir architecture (i.e the same set of hyperparameters as used for test set forecast). It means at each day, the same set of hyperparameter is kept but a new reservoir is generated. Due to the internal random connection of the reservoir, it might result in forecast stochasticity. We observe that models with ridge penalty higher than 1 (i.e all models except multiple IS with input connection) tend to have more stable forecast but that the addition of input layer connection provided some extreme forecasted value. This extreme value phenomena is relatively rare because it is corrected when the number of model for aggregation is above 15.

```{r ucAggregationSpread, echo = FALSE, fig.cap = "Individual forecast of 40 repetition of the best hyperparameter set for the different RC configuration. Forecast value above 200 were set to 200.", fig.height=6}
ls_use_case$plot_before_aggregation_forecast
```

```{r ucAggregationNb, echo = FALSE, fig.cap = "It shows the impact of aggregating several RC with the best set of HP for each RC configuration. It has small interest when the input layer was not connected to output layer but significantly improves the performance otherwise.", fig.height=6}
ls_use_case$plot_aggregation
```

@@

In this specific application, we have demonstrated that RC exhibits commendable performance in comparison to Elastic-net, which serves as the reference model. Furthermore, we highlight the inherent challenges in forecasting within this context, primarily stemming from the non-stationarity of the time series.

This study represents a novel contribution to epidemic forecasting utilizing reservoir computing. Notably, previous literature predominantly focused on simpler problems characterized by fewer input features or shorter evaluation periods. Our findings underscore the potential of this approach for future epidemics, suggesting its potential to surpass more traditional epidemiological tools while maintaining a lightweight model structure compared to other recurrent neural networks (RNNs).

All computations in this study were conducted using the reservoirnet package, and the entire codebase is accessible on GitHub (**todo : open to the public**). The package demonstrates its efficacy in implementing various reservoir architectures, including connection between the input layer and the output layer, as well as the utilization of several input scaling, all within the context of a real-world use case.

Given the substantial number of hyperparameters involved, we acknowledge that random search may not be the most efficient optimization algorithm. We have retained this approach for the sake of simplicity in this tutorial paper; however, meta-heuristic approaches, particularly those utilizing evolutionary algorithms, may prove more efficient, especially when employing multiple input scaling.

It is worth noting that all models, including the @ferte_benefit_2022 models, encounter challenges in accurately anticipating slope shifts, indicating the need for further investigation. Specifically, additional work is warranted to extend the application of RC to high-dimensional settings, building upon the insights gained from models based on a more extensive set of features.

@@

In this paper, we introduce the R package `reservoirnet`, which serves as a versatile tool for implementing reservoir computing based on `ReservoirPy` Python API. It offers flexibility in defining the reservoir architecture, including options for specifying connections between the input layer and the output layer, as well as variations in input scaling as demonstrated on a real-world use case.

We provided a comprehensive overview of the basic usage of the `reservoirnet` package through illustrative examples in regression and classification tasks. This introductory section serves as a foundation for R users, offering step-by-step guidance on constructing and training reservoir computing models using the package. By demonstrating the application of RC in both regression and classification scenarios, we aim to equip users with the essential knowledge and skills needed to harness the capabilities of reservoir computing for diverse tasks.

Drawing on the robust foundation of the `ReservoirPy` structure, a well-maintained library, this package inherits its reliability and longevity. Nevertheless, our emphasis has been on refining the primary functionalities, which may leave more advanced users desiring deeper access to core functions, such as the generation of the reservoir random connections. This initial version of `reservoirnet` must evolve in tandem with the growing understanding and adoption of reservoir computing within the R community.
